Selecting Computations: Theory and Applications

### 2012-03-30 paper/vago.uai/paper.tex
added files, probably too many

\documentclass[]{article}
\usepackage{proceed2e}
\usepackage{enumerate}
\usepackage{algpseudocode}
\usepackage[ruled]{algorithm}
\usepackage{url}
\usepackage{framed}
\usepackage{amsfonts,amsmath,amsthm,amssymb}
\usepackage{graphicx}
\usepackage{url}
\usepackage{color}

@@

\newtheorem{dfn}{Definition}
\newtheorem{thm}{Theorem}
\newtheorem{lmm}{Lemma}
\newtheorem{crl}{Corollary}

@@

\section{Upper Bounds on Value of Information}

The intrinsic (myopic greedy) VOI $\Lambda_i$ of testing an arm $i$ is
the expected reward of the best arm given the information obtained by the test,
minus the expected reward of the best arm according to the current information state.
We will assume here a distribution model where the arm with the highest sample mean
in an information state is the one with the highest expected reward.
Two cases are possible:
\begin{itemize}
\item the arm $\alpha$ with the highest sample mean $\overline
  X_\alpha$ is tested, and $\overline X_\alpha$ becomes lower than
  $\overline X_\beta$ of the second-best arm $\beta$;
\item another arm $i$ is tested, and $\overline X_i$ becomes higher
than $\overline X_\alpha$.
\end{itemize}

In order to estimate the VOI of testing an arm, a sampling model is
needed.  The generic sampling model used is: 
\begin{itemize}
\item The actual value of an arm has a certain (possibly unknown) distribution.
\item Testing an arm is an i.i.d. noisy measurement of the actual
  value of the arm.
\end{itemize}
When the distribution model is known, one can use the methods of Section \ref{sec:normal}.
Otherwise, one can still use {\em concentration
inequalities} to derive distribution-independent bounds on the
VOI. Our bounds are applicable to any bounded distribution. Similar
bounds can be derived for certain unbounded distributions, such as the
normally distributed prior of the value with normally distributed
sampling.

Unfortunately, the pure \textit{myopic} VOI estimate is of little use in
Monte-Carlo sampling, since the effect of a single sample is small,
and the myopic VOI estimate will often be zero. Instead, we will derive
an estimate for the case where we decide to test an arm $N$ times,
where $N$ can be the remaining ``budget'' of available samples or
any other integer quantity.
Denote by  $\Lambda_i^b$ the intrinsic VOI of testing the $i$th arm
$N$ times, and the number of
samples already taken from the $i$th arm by $n_i$.
\begin{thm} $\Lambda_i^b$ is bounded from above as
\begin{align}
\label{eqn:thm-be}
  \Lambda_\alpha^b&\le \frac {N \overline X_\beta^{n_\beta}} {n_\alpha} \Pr(\overline X_\alpha^{n_\alpha+N}\le\overline X_\beta^{n_\beta})\nonumber\\
\Lambda_{i|i\ne\alpha}^b&\le \frac {N(1-\overline X_\alpha^{n_\alpha})} {n_i}\Pr(\overline   X_i^{{n_i}+N}\ge\overline X_\alpha^{n_\alpha})
\end{align}
\label{thm:be}
\end{thm}
\vspace{-2em}
\begin{proof} For the case $i\ne \alpha$, the probability that the
  $i$th arm is finally chosen instead of $\alpha$ is
  $\Pr(\overline X_i^{n_i+N} \ge \overline X_\alpha^{n_\alpha})$. $X_i \le 1$,
  therefore $\overline X_i^{n_i+N}\le \overline
  X_\alpha^{n_\alpha}+\frac {N(1-\overline X_\alpha^{n_\alpha})} {N+n_i}$. Hence, the intrinsic value of perfect
  information is at most: 
\begin{equation}
\label{eq:simplistic}
\frac{ N(1-\overline  X_\alpha^{n_\alpha})}
  {N+n_i}\Pr(\overline X_i^{{n_i}+N}\hspace{-0.5em}\ge\overline X_\alpha^{n_\alpha})\le \frac {N \overline X_\beta^{n_\beta}} {n_\alpha} \Pr(\overline X_\alpha^{n_\alpha+N}\hspace{-0.5em}\le\overline X_\beta^{n_\beta})
\end{equation}
  Proof for the case $i=\alpha$ is similar.
\end{proof}

The probabilities can be bounded from above using the
Hoeffding inequality \cite{Hoeffding.ineq}:
\begin{thm} The probabilities in equations (\ref{eqn:thm-be}) are bounded from above as
\begin{align}
  \label{eqn:probound-blnk-hoeffding}
  \Pr&(\overline X_\alpha^{{n_\alpha}+N} \le \overline X_\beta^{n_\beta})
  \le 2\exp\left(- \varphi (\overline X_\alpha^{n_\alpha} - \overline X_\beta^{n_\beta})^2 n_\alpha
  \right)\nonumber\\
  \Pr&(\overline X_{i|i\ne\alpha}^{n_\alpha+N} \ge \overline X_\beta^{n_\beta})
  \le 2\exp\left(- \varphi (\overline X_\alpha^{n_\alpha} -\overline  X_i^{n_i})^2 n_i \right)
\end{align}
where $\varphi=\min \left(2(\frac {1+n/N} {1+\sqrt {n/N}})^2\right)=8(\sqrt 2 - 1)^2 > 1.37$.
\label{thm:hoeffding-prob-bounds}
\end{thm}
\begin{proof}
Equations (\ref{eqn:probound-blnk-hoeffding}) follow from the
observation that if $i\ne\alpha$, $\overline X_i^{n_i+N}>\overline X_\alpha^{n_i}$
if and only if the mean $\overline X_i^N$ of $N$ samples from $n_i+1$
to $n_i+N$ is at least $\overline X_\alpha^{n_i}+(\overline X_\alpha^{n_i}-\overline
X_i^{n_i})\frac {n_i} N$.

For any $\delta$, the probability that $\overline X_i^{n_i+N}$ is greater
than $\overline X_\alpha^{n_i}$ is less than the probability that
$\IE[X_i]\ge\overline X_i^n+\delta$
\emph{or} $\overline X_i^N\ge \IE[X_i]+\overline X_\alpha^{n_\alpha}
- \overline X_i^{n_i} - \delta +(\overline X_\alpha^{n_\alpha} - \overline X_i^{n_i})\frac {n_i} N$,
thus, by the union bound, less than the sum of the probabilities:
\begin{align}
\Pr&(\overline X_i^{n_i+N}\ge \overline X_\alpha^{n_i})\nonumber\\
   &\le\Pr(\IE[X_i]-\overline X_i^{n_i} \ge \delta)\\
   &\hspace{0.25em}+\Pr\left(\overline X_i^N\hspace{-0.5em} - \IE[X_i] \ge \overline X_\alpha^{n_\alpha}\hspace{-0.5em}
           - \overline X_i^{n_i}\hspace{-0.5em} - \delta +(\overline X_\alpha^{n_\alpha}\hspace{-0.5em} - \overline X_i^{n_i})\frac {n_i} N\right)\nonumber
\end{align}
Bounding the probabilities on the right-hand side using the Hoeffding
inequality, obtain:
\begin{align}
\Pr&(\overline X_i^{n_i+N}\ge \overline X_\alpha^{n_\alpha})\le \nonumber\\
    &\exp(-2\delta^2n_i)+\nonumber\\
    &\quad\exp\left(-2\left((\overline X_\alpha^{n_\alpha}\hspace{-0.5em}
         - \overline X_i^{n_i})\left(1+\frac {n_i} N\right)
         - \delta\right)^2N\right)
\label{eqn:app-hoeffding-le-maxexp}
\end{align}
Find $\delta$ for which the two terms on the right-hand side of
(\ref{eqn:app-hoeffding-le-maxexp}) are equal:
\begin{equation}
\exp(-\delta^2n) = \exp\left(-2\left((\overline X_\alpha - \overline X_i)(1+\frac {n_i} N) - \delta\right)^2N\right)\label{eqn:app-hoeffding-eq-exp}
\end{equation}
Solve (\ref{eqn:app-hoeffding-eq-exp}) for $\delta$: $\delta=\frac {1+\frac {n_i} N} {1+\sqrt {\frac {n_i} N}} (\overline X_\alpha^{n_\alpha}\hspace{-0.5em}
- \overline X_i^{n_i}) \ge 2(\sqrt 2 - 1)(\overline X_\alpha^{n_\alpha}\hspace{-0.5em}-\overline X_i^{n_i})$. Substitute $\delta$ into 
(\ref{eqn:app-hoeffding-le-maxexp}) and obtain
\begin{align}
\Pr&(\overline X_i^{n_i}\ge \overline X_\alpha^{n_\alpha}) \nonumber\\
& \le 2\exp\left(-2\left( \frac {1+\frac {n_i} N} {1+\sqrt {\frac {n_i} N}}
                          (\overline X_\alpha^{n_\alpha} - \overline X_i^{n_i})\right)^2 n_i\right)\nonumber \\
& \le 2\exp(-8(\sqrt 2 - 1)^2(\overline X_\alpha^{n_\alpha} - \overline X_i^{n_i})^2n_i)\nonumber\\
& = 2\exp(-\varphi(\overline X_\alpha^{n_\alpha} - \overline X_i^{n_i})^2n_i)
\end{align}
Derivation for the case $i=\alpha$ is similar.
\end{proof}

@@

More refined bounds can be obtained through tighter estimates on the
probabilities in Equation \ref{eq:simplistic}, for example, based on the empirical Bernstein
inequality~\cite{MaurerPontil.benrstein}, or through a more careful
application of the Hoeffding inequality, resulting in:
\begin{align}
\Lambda_i^b&\le\frac {N\sqrt \pi} {n_i \sqrt {n_i}}
  \left[\mathrm{erf}\left((1\hspace{-0.25em}-\hspace{-0.25em}\overline X_i^{n_i})\sqrt {n_i}\right)
      \hspace{-0.25em}-\hspace{-0.25em}\mathrm{erf}\left((\overline X_\alpha^{n_\alpha}\hspace{-0.5em} - \overline X_i^{n_i})\sqrt{n_i}\right)\right]\nonumber\\
\Lambda_\alpha^b&\le\frac {N\sqrt \pi} {n_\alpha \sqrt {n_\alpha}}
  \left[\mathrm{erf}\left(\overline X_\alpha^{n_\alpha}\sqrt {n_\alpha}\right)
      -\mathrm{erf}\left((\overline X_\alpha^{n_\alpha}\hspace{-0.5em} - \overline X_\beta^{n_\beta})\sqrt{n_\alpha}\right)\right]
\label{eqn:erf-blinkered}
\end{align}

\section{VOI-based Sampling Control}

\subsection{Selection Criterion}

A reasonable greedy policy is to test the arm that has the best
current VOI estimate. When the distributions are unknown, it makes sense
to use the upper bounds established in Theorem~\ref{thm:be} instead. Although MCTS
differs from pure exploration in Multi-armed Bandits in that the
distributions of the rewards are not stationary, VOI
estimates computed as for stationary distributions work well in
practice. As illustrated
by the empirical evaluation (Section~\ref{sec:empirical-evaluation}),
estimates based on upper bounds on the VOI result in good sampling
policies, which exhibit performance comparable to the performance of
some state-of-the-art heuristic algorithms.

\subsection{Stopping Criterion}
\label{sec:control-stopping-criterion}

The simplest stopping criterion for a sampling policy is the
budget---a fixed number of samples performed before a decision is
made. When a sample has a known cost commensurable with the value of
information of a measurement, an upper bound on the intrinsic VOI can also
be used to stop the sampling if the intrinsic VOI of any action
is less than the cost of sampling $C$:
\begin{equation}
\mbox{stop if } \max_i \Lambda_i \le C
\end{equation}
VOI estimates~(\ref{eqn:thm-be},
\ref{eqn:bound-blnk-hoeffding}) include the remaining sample budget $N$ as a
factor, but given the cost of a single sample $c$, the cost of the
remaining samples accounted for in estimating the intrinsic VOI is
$C=cN$. $N$ can be dropped on both sides of the inequality,
and a reasonable stopping criterion would be:
\begin{align}
\frac 1 N \Lambda_\alpha^b \le&\frac {\overline X_\beta^{n_\beta}}
  {n_\alpha}\Pr(\overline X_\alpha^{n_\alpha+N}\le\overline
  X_\beta^{n_\alpha})\le c\nonumber\\
\frac 1 N \max_i\Lambda_i^b\le &\max_i\frac {(1-\overline X_\alpha^{n_\alpha})} {n_i}\Pr(\overline
  X_i^{n_i+N}\ge\overline X_\alpha^{n_\alpha})\le c\nonumber\\
    &\forall i: i\ne\alpha
\label{eqn:stopping-blnk}
\end{align}
The empirical evaluation (Section~\ref{sec:empirical-evaluation})
confirms the viability of this stopping criterion and illustrates the
influence of the sample cost $c$ on the performance of
the sampling policy.

\subsection{Sample Redistribution in Trees}
\label{sec:control-redistribution}

Monte-Carlo tree search \cite{Chaslot.montecarlo} selects at each
search step an action that appears to be the best based on outcomes
of \textit{search rollouts}. Two different criteria are employed in
action selection during a rollout:
\begin{itemize}
\item at the first step, the value of partial
information should be used to select nodes to sample, as the goal is
selection of a move at the current root node;
\item as the goal of sampling in deeper tree nodes is usually to improve the
node value estimate, rather than selection, a different sampling scheme
should be used there. At these levels, the standard practice
of using UCT-like rollouts makes sense, as this method is known to be nearly optimal
in minimizing the \textit{cumulative} regret.
\end{itemize}

The above hybrid approach assumes
that the information obtained from rollouts in the
current state is discarded after an action is selected. In practice,
many successful Monte-Carlo tree search algorithms re-use rollouts
generated at earlier search states, if the sample traverses the
current search state during the rollout; thus, the value of information of a rollout is
determined not just by the influence on the choice of the action at
the current state, but also by its potential influence on the choice at future
search states.

@@

\section{Empirical Evaluation}
\label{sec:empirical-evaluation}

The sampling policies are first compared on random Multi-armed bandit 
problem instances. Figure~\ref{fig:random-instances} shows results for
randomly-generated Multi-armed bandits with 25 Bernoulli arms, with
the mean rewards of the arms distributed uniformly in the range~$[0,
  1]$, for a range of sample budgets~$200..2000$, with multiplicative
step of~$2$. The experiment for each number of samples was repeated
10000 times. UCB1, as well as policies  based on bounds
(\ref{eqn:bound-blnk-hoeffding}) (VOI) and
(\ref{eqn:erf-blinkered}) (VOI+) are compared. UCB1 is always considerably worse than the
VOI-aware sampling policies.

\subsection{Playing Go Against UCT}
\label{sec:emp-go}

>>>>

\documentclass[]{article}
\usepackage{proceed2e}
\usepackage{enumerate}
\usepackage{algpseudocode}
\usepackage[ruled]{algorithm}
\usepackage{url}
\usepackage{framed}
\usepackage{amsfonts,amsmath,amsthm,amssymb}
\usepackage{graphicx}
\usepackage{url}
\usepackage{color}
\usepackage[round]{natbib}
\usepackage{comment}

\def\citeapos#1{\citeauthor{#1}'s (\citeyear{#1})}

@@

\newcommand {\term}[1] {\textbf{#1}}
\newcommand {\note}[1] {\textcolor{red}{[[#1]]}}
\newcommand {\Evidence} {\mathcal{E}}
\newcommand {\R} {\mathbb{R}}
\newcommand {\maps} {\colon}
\newcommand {\given} {\mathrel{|}}
\newcommand {\sample} {\sim}
\newcommand {\iid} {\stackrel{\rm iid}{\sample}}
\DeclareMathOperator*{\argmax}{argmax}

\newtheorem{thm}{Theorem}
\newtheorem{dfn}[thm]{Definition}
\newtheorem{lmm}[thm]{Lemma}
\newtheorem{crl}[thm]{Corollary}
\newtheorem{example}{Example}

\newcommand{\secref}[1]{Section~\ref{#1}}
\newcommand{\secrefs}[2]{Sections~\ref{#1} and~\ref{#2}}
\newcommand{\figref}[1]{Figure~\ref{#1}}
\renewcommand{\eqref}[1]{Equation~(\ref{#1})}
\newcommand{\eqrefs}[2]{Equations~(\ref{#1}) and~(\ref{#2})}
\newcommand{\thmref}[1]{Theorem~\ref{#1}}
\newcommand{\dfnref}[1]{Definition~\ref{#1}}
\newcommand{\exampleref}[1]{Example~\ref{#1}}

\title{Selecting Computations: Theory and Applications}

\author{}
%\author{ {\bf Nicholas Hay} and {\bf Stuart Russell} \\   
%Computer Science Division \\  
%University of California\\ 
%Berkeley, CA 94720 \\ 
%\And 
%{\bf Solomon Eyal Shimony} and {\bf David Tolpin}  \\ 
%Department of Computer Science \\ 
%Ben-Gurion University of the Negev\\
%Beer Sheva, Israel
%}

@@

\maketitle

\begin{abstract}
Sequential decision problems are often approximately solvable by
simulating possible future action sequences.  {\em Metalevel} decision
procedures have been developed for selecting {\em which} action
sequences to simulate, based on some estimate of the expected
improvement in decision quality resulting from any particular
simulation; an example is the recent work on using selection rules
derived for bandit problems to control Monte Carlo tree search in the
game of Go.  In this paper we develop a theoretical basis for
metalevel decisions in the statistical framework of {\em selection problems}, 
arguing that this is more appropriate than the bandit
framework.  We derive a number of basic results applicable to simple
Monte Carlo selection problems, including the first finite sampling
bounds for optimal policies in certain cases; we also provide a simple
counterexample to the intuitive conjecture that an optimal policy will
necessarily reach a decision in all cases.  We then derive heuristic
approximations in both Bayesian and distribution-free settings and
demonstrate their superiority to bandit-based heuristics in one-shot
decision problems and in Go.
\end{abstract}

\section{Introduction}

\input{introduction}

%% \input{related}

\section{On optimal policies for selection}\label{sec:optimal}

\input{optimalpolicies}

\section{Context effects and non-indexability}\label{sec:context}

\input{context}

\section{The blinkered policy}\label{sec:blinkered}\label{approx-bayesian-section}

\input{blinkered}

%% \section{Regret bounds and approximate policies}

%% [[Regret models: simple regret, regret with cost per sampling; regret goes to zero as c does]]

%% [[Expected simple regret bounds for normal case?]]

%% [[Blinkered sampling]]

%% [[ESPb: Frazier's continuous time approximation]]

\section{Upper bounds on value of information}\label{approx-nonbayesian-section}

%% Solomon: Text presumably should be placed after the relevant theorems.

Theorem \ref{thm:optimal-myopic} provides necessary conditions
for a stopping condition for the optimal policy in terms of a myopic
policy, but that does not preclude premature stopping of a greedy policy.
Conversely, Theorem \ref{thm:myopic-optimal} provides sufficient conditions, but these 
conditions are hard to evaluate in practice.

The myopic selection policy uses
the intrinsic value of information (VOI) $\Lambda_i$ of testing an arm $i$, which is
the expected reward of the best arm given the information obtained by the test,
minus the expected reward of the best arm according to the current information state.
Unfortunately, the pure \textit{myopic} VOI estimate is of little use in
Monte-Carlo sampling, since the effect of a single sample is small,
and the myopic VOI estimate will often be zero.

Instead, we generalize myopic policies to include ``semi-myopic"
up to $N$ samples of the same arm; these are called ``blinkered" policies.
Such policies allow potentially unlimited lookahead, but only in a single ``direction'' (one specific arm),
as if we ``had our blinkers on''.

%% Solomon: end text to be (possibly moved).

In many practical applications of the selection problem, such as search in
the game of Go, prior distributions are unavailable\footnote{The analysis is also applicable to
some Bayesian settings, using ``fake" samples to simulate prior distributions.}
In such cases, one can still bound
the value of information w.r.t. myopic policies using {\em concentration
inequalities} to derive distribution-independent bounds on the
VOI. We obtain such bounds under the
following assumptions:
\begin{enumerate}
\item Samples are iid given the value of the arms (variables), as in the Bayesian schemes such as Bernoulli
sampling.
\item The expectation of a selection in a belief state is equal to the sample mean (and therefore,
   after sampling terminates, the arm with the greatest sample mean will be selected).
\end{enumerate}

When considering possible samples in the ``blinkered'' semi-myopic setting,
two cases are possible:
\begin{itemize}
\item the arm $\alpha$ with the highest sample mean $\overline
  X_\alpha$ is tested, and $\overline X_\alpha$ becomes lower than
  $\overline X_\beta$ of the second-best arm $\beta$;
\item another arm $i$ is tested, and $\overline X_i$ becomes higher
than $\overline X_\alpha$.
\end{itemize}

Our bounds below are applicable to any bounded distribution (with rewards being
bounded to $[0,1]$). Similar
bounds can be derived for certain unbounded distributions, such as the
normally distributed prior of the value with normally distributed
sampling.
We derive a VOI bound for testing an arm $N$ times,
where $N$ can be the remaining ``budget'' of available samples or
any other integer quantity.
Denote by  $\Lambda_i^b$ the intrinsic VOI of testing the $i$th arm
$N$ times, and the number of
samples already taken from the $i$th arm by $n_i$.
\begin{thm} $\Lambda_i^b$ is bounded from above as
\begin{align}
\label{eqn:thm-be}
  \Lambda_\alpha^b&\le \frac {N \overline X_\beta^{n_\beta}} {n_\alpha} \Pr(\overline X_\alpha^{n_\alpha+N}\le\overline X_\beta^{n_\beta})\nonumber\\
\Lambda_{i|i\ne\alpha}^b&\le \frac {N(1-\overline X_\alpha^{n_\alpha})} {n_i}\Pr(\overline   X_i^{{n_i}+N}\ge\overline X_\alpha^{n_\alpha})
\end{align}
\label{thm:be}
\end{thm}
\vspace{-2em}
\begin{proof} For the case $i\ne \alpha$, the probability that the
  $i$th arm is finally chosen instead of $\alpha$ is
  $\Pr(\overline X_i^{n_i+N} \ge \overline X_\alpha^{n_\alpha})$. $X_i \le 1$,
  therefore $\overline X_i^{n_i+N}\le \overline
  X_\alpha^{n_\alpha}+\frac {N(1-\overline X_\alpha^{n_\alpha})} {N+n_i}$. Hence, the intrinsic value of blinkered
  information is at most: 
\begin{align}
\label{eq:simplistic}
\frac{ N(1-\overline  X_\alpha^{n_\alpha})}
  {N+n_i}&\Pr(\overline X_i^{{n_i}+N}\hspace{-0.5em}\ge\overline X_\alpha^{n_\alpha})\nonumber \\
&\le\frac{ N(1-\overline  X_\alpha^{n_\alpha})}
{n_i}\Pr(\overline X_i^{{n_i}+N}\hspace{-0.5em}\ge\overline X_\alpha^{n_\alpha})
\end{align}
  Proof for the case $i=\alpha$ is similar.
\end{proof}

The probabilities can be bounded from above using the
Hoeffding inequality \citep{Hoeffding.ineq}:
\begin{thm} The probabilities in equations (\ref{eqn:thm-be}) are bounded from above as
\begin{align}
  \label{eqn:probound-blnk-hoeffding}
  \Pr&(\overline X_\alpha^{{n_\alpha}+N} \le \overline X_\beta^{n_\beta})
  \le 2\exp\left(- \varphi (\overline X_\alpha^{n_\alpha} - \overline X_\beta^{n_\beta})^2 n_\alpha
  \right)\nonumber\\
  \Pr&(\overline X_{i|i\ne\alpha}^{n_\alpha+N} \ge \overline X_\beta^{n_\beta})
  \le 2\exp\left(- \varphi (\overline X_\alpha^{n_\alpha} -\overline  X_i^{n_i})^2 n_i \right)
\end{align}
where $\varphi=\min \left(2(\frac {1+n/N} {1+\sqrt {n/N}})^2\right)=8(\sqrt 2 - 1)^2 > 1.37$.
\label{thm:hoeffding-prob-bounds}
\end{thm}

\begin{comment}
	\begin{proof}
	Equations (\ref{eqn:probound-blnk-hoeffding}) follow from the
	observation that if $i\ne\alpha$, $\overline X_i^{n_i+N}>\overline X_\alpha^{n_i}$
	if and only if the mean $\overline X_i^N$ of $N$ samples from $n_i+1$
	to $n_i+N$ is at least $\overline X_\alpha^{n_i}+(\overline X_\alpha^{n_i}-\overline
	X_i^{n_i})\frac {n_i} N$.

	For any $\delta$, the probability that $\overline X_i^{n_i+N}$ is greater
	than $\overline X_\alpha^{n_i}$ is less than the probability that
	$\IE[X_i]\ge\overline X_i^n+\delta$
	\emph{or} $\overline X_i^N\ge \IE[X_i]+\overline X_\alpha^{n_\alpha}
	- \overline X_i^{n_i} - \delta +(\overline X_\alpha^{n_\alpha} - \overline X_i^{n_i})\frac {n_i} N$,
	thus, by the union bound, less than the sum of the probabilities:
	\begin{align}
	\Pr&(\overline X_i^{n_i+N}\ge \overline X_\alpha^{n_i})\nonumber\\
	   &\le\Pr(\IE[X_i]-\overline X_i^{n_i} \ge \delta)\\
	   &\hspace{0.25em}+\Pr\left(\overline X_i^N\hspace{-0.5em} - \IE[X_i] \ge \overline X_\alpha^{n_\alpha}\hspace{-0.5em}
	           - \overline X_i^{n_i}\hspace{-0.5em} - \delta +(\overline X_\alpha^{n_\alpha}\hspace{-0.5em} - \overline X_i^{n_i})\frac {n_i} N\right)\nonumber
	\end{align}
	Bounding the probabilities on the right-hand side using the Hoeffding
	inequality, obtain:
	\begin{align}
	\Pr&(\overline X_i^{n_i+N}\ge \overline X_\alpha^{n_\alpha})\le \nonumber\\
	    &\exp(-2\delta^2n_i)+\nonumber\\
	    &\quad\exp\left(-2\left((\overline X_\alpha^{n_\alpha}\hspace{-0.5em}
	         - \overline X_i^{n_i})\left(1+\frac {n_i} N\right)
	         - \delta\right)^2N\right)
	\label{eqn:app-hoeffding-le-maxexp}
	\end{align}
	Find $\delta$ for which the two terms on the right-hand side of
	(\ref{eqn:app-hoeffding-le-maxexp}) are equal:
	\begin{equation}
	\exp(-\delta^2n) = \exp\left(-2\left((\overline X_\alpha - \overline X_i)(1+\frac {n_i} N) - \delta\right)^2N\right)\label{eqn:app-hoeffding-eq-exp}
	\end{equation}
	Solve (\ref{eqn:app-hoeffding-eq-exp}) for $\delta$: $\delta=\frac {1+\frac {n_i} N} {1+\sqrt {\frac {n_i} N}} (\overline X_\alpha^{n_\alpha}\hspace{-0.5em}
	- \overline X_i^{n_i}) \ge 2(\sqrt 2 - 1)(\overline X_\alpha^{n_\alpha}\hspace{-0.5em}-\overline X_i^{n_i})$. Substitute $\delta$ into 
	(\ref{eqn:app-hoeffding-le-maxexp}) and obtain
	\begin{align}
	\Pr&(\overline X_i^{n_i}\ge \overline X_\alpha^{n_\alpha}) \nonumber\\
	& \le 2\exp\left(-2\left( \frac {1+\frac {n_i} N} {1+\sqrt {\frac {n_i} N}}
	                          (\overline X_\alpha^{n_\alpha} - \overline X_i^{n_i})\right)^2 n_i\right)\nonumber \\
	& \le 2\exp(-8(\sqrt 2 - 1)^2(\overline X_\alpha^{n_\alpha} - \overline X_i^{n_i})^2n_i)\nonumber\\
	& = 2\exp(-\varphi(\overline X_\alpha^{n_\alpha} - \overline X_i^{n_i})^2n_i)
	\end{align}
	Derivation for the case $i=\alpha$ is similar.
	\end{proof}	
\end{comment}

@@

More refined bounds can be obtained through tighter estimates on the
probabilities in Equation \ref{eq:simplistic}, for example, based on the empirical Bernstein
inequality~\citep{MaurerPontil.benrstein}, or through a more careful
application of the Hoeffding inequality, resulting in:
\begin{align}
\Lambda_i^b&\le\frac {N\sqrt \pi} {n_i \sqrt {n_i}}
  \left[\mathrm{erf}\left((1\hspace{-0.25em}-\hspace{-0.25em}\overline X_i^{n_i})\sqrt {n_i}\right)
      \hspace{-0.25em}-\hspace{-0.25em}\mathrm{erf}\left((\overline X_\alpha^{n_\alpha}\hspace{-0.5em} - \overline X_i^{n_i})\sqrt{n_i}\right)\right]\nonumber\\
\Lambda_\alpha^b&\le\frac {N\sqrt \pi} {n_\alpha \sqrt {n_\alpha}}
  \left[\mathrm{erf}\left(\overline X_\alpha^{n_\alpha}\sqrt {n_\alpha}\right)
      -\mathrm{erf}\left((\overline X_\alpha^{n_\alpha}\hspace{-0.5em} - \overline X_\beta^{n_\beta})\sqrt{n_\alpha}\right)\right]
\label{eqn:erf-blinkered}
\end{align}

\section{Practical sampling control: empirical evaluation}
\label{sec:empirical-evaluation}\label{mcts-section}

As the selection MDP is intractable, we use myopic or semi-myopic policies instead.
The myopic policies usually separate out the decision of {\em whether} to
sample or to stop (called the stopping policy), and {\em what} to sample.
We first examine the latter issue.

Assuming that the sample costs are constant,
a semi-myopic policy will decide to test the arm that has the best
current VOI estimate. 
When the distributions are unknown, it makes sense
to use the upper bounds established in Theorem~\ref{thm:be} instead,
which is the method adopted here.
This evaluation assumes a fixed budget of samples, which is
completely used up by each of the candidate schemes, making a stopping
criterion irrelevant.

The sampling policies are compared on random Multi-armed bandit 
selection problem instances. Figure~\ref{fig:random-instances} shows results for
randomly-generated Multi-armed bandits with 25 Bernoulli arms, with
the mean rewards of the arms distributed uniformly in the range~$[0,
  1]$, for a range of sample budgets~$200..2000$, with multiplicative
step of~$2$. The experiment for each number of samples was repeated
10000 times. UCB1, as well as policies  based on bounds
(\ref{eqn:bound-blnk-hoeffding}) (VOI) and
(\ref{eqn:erf-blinkered}) (VOI+) are compared. UCB1 is always considerably worse than the
VOI-aware sampling policies.

\subsection{Sampling in trees}

Sampling in trees is considerably more complicated.
The goal of Monte-Carlo tree search \citep{Chaslot.montecarlo} at the root node 
is usually to select an action that appears to be the best based on outcomes
of \textit{search rollouts}.
But the goal of rollouts at non-root nodes
is different than at the root: here it is important to better approximate the ``true'' 
value of the node, so that selection at the root can be more informed. The exact analysis
of sampling at internal nodes is outside the scope of this paper. At present we 
have no better proposal for internal nodes than to use UCT there.

We thus propose the following hybrid sampling scheme:
\begin{itemize}
\item at the root node, sample based on the VOI estimate;
\item at non-root nodes, sample using UCT.
\end{itemize}

Strictly speaking, even at the root node the assumptions underlying our belief-state
MDP for selection do not hold exactly. UCT is an adaptive scheme, and therefore the values
generated by sampling at non-root nodes will typically cause valued observed at
children of the root node to be non-stationary. 
Nevertheless, sampling based on VOI estimates
computed as for stationary distributions work well in
practice. As illustrated
by the empirical evaluation (Section~\ref{sec:empirical-evaluation}),
estimates based on upper bounds on the VOI result in good sampling
policies, which exhibit performance comparable to the performance of
some state-of-the-art heuristic algorithms.

\subsubsection{Stopping criterion}
\label{sec:control-stopping-criterion}

When a sample has a known cost commensurable with the value of
information of a measurement, an upper bound on the intrinsic VOI can also
be used to stop the sampling if the intrinsic VOI of any action
is less than the cost of sampling $C$:
\begin{equation}
\mbox{stop if } \max_i \Lambda_i \le C
\end{equation}
VOI estimates~(\ref{eqn:thm-be},
\ref{eqn:bound-blnk-hoeffding}) include the remaining sample budget $N$ as a
factor, but given the cost of a single sample $c$, the cost of the
remaining samples accounted for in estimating the intrinsic VOI is
$C=cN$. $N$ can be dropped on both sides of the inequality,
and a reasonable stopping criterion would be:
\begin{align}
\frac 1 N \Lambda_\alpha^b \le&\frac {\overline X_\beta^{n_\beta}}
  {n_\alpha}\Pr(\overline X_\alpha^{n_\alpha+N}\le\overline
  X_\beta^{n_\alpha})\le c\nonumber\\
\frac 1 N \max_i\Lambda_i^b\le &\max_i\frac {(1-\overline X_\alpha^{n_\alpha})} {n_i}\Pr(\overline
  X_i^{n_i+N}\ge\overline X_\alpha^{n_\alpha})\le c\nonumber\\
    &\forall i: i\ne\alpha
\label{eqn:stopping-blnk}
\end{align}
The empirical evaluation (Section~\ref{sec:empirical-evaluation})
confirms the viability of this stopping criterion and illustrates the
influence of the sample cost $c$ on the performance of
the sampling policy. When the sample cost $c$ is unknown, one can perform initial calibration experiments
to determine a reasonable value, as done in this paper.

\subsubsection{Sample redistribution in trees}
\label{sec:control-redistribution}

The above hybrid approach assumes
that the information obtained from rollouts in the
current state is discarded after an real-world action is selected. In practice,
many successful Monte-Carlo tree search algorithms re-use rollouts
generated at earlier search states, if the sample traverses the
current search state during the rollout; thus, the value of information of a rollout is
determined not just by the influence on the choice of the action at
the current state, but also by its potential influence on the choice at future
search states.

@@

\subsubsection{Playing Go against UCT}
\label{sec:emp-go}

The hybrid policies were compared on Computer Go, a search domain
in which UCT-based MCTS has been particularly successful
\citep{Gelly.mogo}. A modified version of Pachi \citep{Braudis.pachi}, a state of the art
Go program, was used for the experiments:
\begin{itemize}
\item The UCT engine of Pachi was extended with VOI-aware sampling
  policies at the first step. 
\item The stopping criterion for the VOI-aware policy was
  modified and based solely on the sample cost, specified as
  a constant parameter. The heuristic stopping criterion for the
  original UCT policy was left unchanged.
\item The time-allocation mode based on the fixed number of samples
  was modified for \textit{both the original UCT policy and the VOI-aware
  policies} such that 
  \begin{itemize}
    \item the same number of samples is available to
      the agent at each step, independently of the number of pre-simulated
      games;  
    \item if samples were unused at the current step,
      they become available at the next step. 
  \end{itemize}
\end{itemize}
While the UCT engine is not the most powerful engine of Pachi, it is still a strong
player. On the other hand, additional features of more advanced
engines would obstruct the MCTS phenomena which are the subject of
the experiment.
\begin{figure}[h!]
\centering
\includegraphics[scale=0.6]{uctvoi.pdf}
\caption{Go: winning rate vs. $c$}
\label{fig:uctvoi}
\end{figure}
The engines were compared on the 9x9 board, for 5000, 7000, 1000, and
15000 samples (game simulations) per ply, each experiment was repeated
1000 times. Figure~\ref{fig:uctvoi} depicts a calibration experiment,
showing the winning rate of the VOI-aware policy against UCT
vs. the stopping threshold $c$ (if the maximum VOI of a sample is below
the threshold, the simulation is stopped, and a move is chosen). Each
curve in the figure corresponds to a certain number of samples per
ply. For the stopping threshold of $10^{-6}$, the VOI-aware policy
is almost always better than UCT, and  reaches the winning rate of
64\% for 10000 samples per ply.
